---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# rtide 

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check](https://github.com/millerlp/rtide/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/millerlp/rtide/actions/workflows/R-CMD-check.yaml)
<!--[![Codecov test coverage](https://codecov.io/gh/poissonconsulting/rtide/graph/badge.svg)](https://app.codecov.io/gh/poissonconsulting/rtide) --> 
[![License: GPL3](https://img.shields.io/badge/License-GPL3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
[![CRAN status](https://www.r-pkg.org/badges/version/rtide)](https://cran.r-project.org/package=rtide)
![CRAN downloads](https://cranlogs.r-pkg.org/badges/rtide)
[![Codecov test coverage](https://codecov.io/gh/millerlp/rtide/graph/badge.svg)](https://app.codecov.io/gh/millerlp/rtide)
<!-- badges: end -->

## Introduction

`rtide` is an R package to calculate tide heights based on tide station harmonics.

It includes the harmonics data for `r length(rtide::tide_stations())` US stations.

## Installation

To install the latest release from [CRAN](https://cran.r-project.org)
```{r, eval=FALSE}
install.packages("rtide")
```

To install the developmental version from [GitHub](https://github.com/millerlp/rtide)
```{r, eval=FALSE}
# install.packages("pak")
pak::pak("millerlp/rtide")
```

## Utilisation

```{r}
library(tibble)
library(rtide)

data <- rtide::tide_height(
  "Monterey Harbor",
  from = as.Date("2016-07-13"), to = as.Date("2016-07-15"),
  minutes = 10L, tz = "America/Los_Angeles"
)

print(data)
```

```{r, message=FALSE}
library(ggplot2)
library(scales)
```

```{r}
ggplot(data = data, aes(x = DateTime, y = TideHeight)) +
  geom_line() +
  scale_x_datetime(
    name = "Date",
    labels = date_format("%d %b %Y", tz = "America/Los_Angeles")
  ) +
  scale_y_continuous(name = "Tide Height (m)") +
  ggtitle("Monterey Harbor")
```

## Shiny

Tide heights can be also obtained using rtide through a [shiny interface](https://poissonconsulting.shinyapps.io/rtide/) developed by Seb Dalgarno.

## Contribution

Please report any [issues](https://github.com/millerlp/rtide/issues).

[Pull requests](https://github.com/millerlp/rtide/pulls) are always welcome.

<!--  ## Code of Conduct
  
Please note that the rtide project is released with a [Contributor Code of Conduct](https://poissonconsulting.github.io/rtide/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms. -->

## Inspiration

The harmonics data was converted from <https://github.com/millerlp/rtide/blob/main/data-raw/harmonics-dwf-20251228-free.tar.xz>, NOAA web site data
processed by David Flater for [`XTide`](https://flaterco.com/xtide/).
The code to calculate tide heights from the harmonics is based on `XTide`.

## Tide prediction updates

The harmonic constituent data for the various tide stations are updated yearly 
by the XTide maintainer. The `rtide` package originally used a harmonic data base
from 2015. An updated harmonic constituent database file from 2025-12-28 has 
been incorporated into `rtide` as of v0.0.12.9000 (August 2026), which adds 
many new tide stations, as well as slightly changing the harmonic constituent 
values for many stations. As a result, tide height predictions produced with 
earlier versions of the `rtide` package (v0.0.12 and earlier) may produce 
slightly different values than the predictions generated by the new harmonic 
data base. Those differences may typically only amount to a centimeter or less
compared to previous predictions generated with earlier versions of the harmonic
constituents for a given site. 
