---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![Travis-CI Build Status](https://travis-ci.org/poissonconsulting/rtide.svg?branch=master)](https://travis-ci.org/poissonconsulting/rtide)
[![codecov](https://codecov.io/gh/poissonconsulting/rtide/branch/master/graph/badge.svg)](https://codecov.io/gh/poissonconsulting/rtide)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rtide)](https://cran.r-project.org/package=rtide)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/grand-total/rtide)](https://cran.r-project.org/package=rtide)

# rtide

## Introduction

`rtide` is an R package to calculate tide heights based on tide station harmonics.

It includes the harmonics data for `r length(rtide::tide_stations())` US stations.

## Utilisation

```{r}
library(rtide)

data <- rtide::tide_height(
  "Monterey Harbor", from = as.Date("2016-07-13"), to = as.Date("2016-07-15"), 
  minutes = 10L, tz = "PST8PDT")

print(head(data))
```

```{r, message=FALSE}
library(ggplot2)
library(scales)
```

```{r}
ggplot(data = data, aes(x = DateTime, y = TideHeight)) + 
  geom_line() + 
  scale_x_datetime(name = "Date", 
                   labels = date_format("%d %b %Y", tz="PST8PDT")) +
  scale_y_continuous(name = "Tide Height (m)") +
  ggtitle("Monterey Harbour")
```

## Installation

To install the release version from CRAN
```
install.packages("rtide")
```

Or the development version from GitHub
```
# install.packages("devtools")
devtools::install_github("poissonconsulting/rtide")
```
## Citation

```{r, comment="", echo=FALSE}
citation(package = "rtide")
```

## Interaction

Tide heights can be also obtained using rtide through a [shiny interface](https://poissonconsulting.shinyapps.io/rtide/) developed by Seb Dalgarno.

## Contribution

Please report any [issues](https://github.com/poissonconsulting/rtide/issues).

[Pull requests](https://github.com/poissonconsulting/rtide/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.

## Inspiration

The harmonics data was converted from harmonics-dwf-20151227-free, NOAA web site data
processed by David Flater for [XTide](http://www.flaterco.com/xtide/).
The code to calculate tide heights from the harmonics is based on XTide.
